Eleven years of student replication projects provide evidence on the correlates of replicability in psychology

### 2023-07-10 paper/manuscript.Rmd
new draft

Cumulative scientific progress requires empirical results that are robust enough to support theory construction and extension. Yet in psychology, some prominent findings have failed to replicate, and large-scale studies suggest replicability issues are widespread. The identification of predictors of replication success is limited by the difficulty of conducting large samples of indepedendent replication experiments, however: most investigations re-analyse the same set of ~170 replications. We introduce a new dataset of 176 replications from students in a graduate-level methods course. Replication results were judged to be successful in 49% of replications; of the 136 where comparable effect sizes could be computed, 45% had point estimates within the prediction interval of the original outcome. Larger original effect sizes and within-participants designs were especially related to replication success. Or results indicate that, consistent with prior reports, the robustness of the psychology literature is low enough to limit cumulative progress by student investigators.
:::

@@

A few large-scale investigations have measured replication rates in samples of psychology studies. The first of these, RP:P, sampled roughly 100 studies from articles published in three top psychology journals in 2008 and distributed the studies across participating labs, finding an overall replication rate of around 40% [@openscienceconsortium2015]. The followup Many Labs Studies investigated heterogeneity using short target studies that each compared two conditions. These study designs were not representative of the psychology literature as a whole, and due to the of measuring heterogeneity, they had large overall samples across multiple sites. Across Many Labs 1--3, only 29 of 51 target effects (57%) replicated [@klein2014; @klein2018; @ebersole2016]. @camerer2018 included all 21 behavioral social science studies from Nature and Science from 2010-2015 that were feasible to replicate. They consulted original authors and had high power to detect even small effects; under these conditions and with this sample, the replication rate was around 60%.

@@

Further, the methods used in these studies are not sufficient to yield an unbiased estimate of this quantity. In no case were studies randomly sampled from the literature; instead replication projects sampled from specific journals and adjusted the sample for feasibility concerns. These reasonable decisions further undermine interpretation of the results from these studies as representing the proportion of true findings in the psychology literature as a whole. 
<!-- Sampling at random from the literature may not even be desirable, as arguably a literature will succeed if useful discoveries come out of it, not if random findings are true [@wilson2020].  -->

Rather than aim for some measure of true replicability, perhaps we should contextualize the true estimand for replication efforts based on their methodologies and outcome measures. Through this lens, RP:P estimated the rate at which findings from relatively simple experiments published in a few well-known journals at a particular time could be replicated in a typical psychology lab. Many Labs estimated at which well-known, two-condition findings replicate in very large samples. @camerer2018 estimated which prestigious journal findings replicate when conducted in a highly-resourced environment with expert involvement. All of these could be potentially desirable estimands.

But in practice, most scientific work is conducted by graduate students with limited time, limited budgets and limited access to experts. If students cannot replicate a finding under these circumstances, they cannot build on it in their empirical or theoretical work. How replicable are findings in the literature for graduate students operating under these less-than-ideal conditions? We address this question here.

@@

In search of such features, RP:P examined correlates of replicability in the RP:P sample and found that studies in cognitive psychology (as opposed to social psychology) and studies with larger effect sizes and smaller p-values were more likely to replicate. Using these same data combined with a few other smaller samples, @altmejd2019 examined statistical and demographic features of replication studies and identified larger sample sizes, larger effect sizes, and simple effects (as opposed to interaction terms) as predictive of replication.^[While most approaches to correlates of replicability have been correlational, experimental approaches can be used to test potential interventions [@protzko2020;@ebersole2020]. Experiments can be very valuable as tests of specific causes of non-replication, but they are expensive and time-consuming to conduct, and not all factors affecting replication success can be manipulated experimentally.]
<!-- showed, across 16 studies, that better methodological practices led to replication rates that matched theoretical expectations based on sampling error, with replication effect sizes comparable with the original. Many Labs 5 added expert (original author) advice to a replication process, and found that it did not substantially increase the replication rate for the 10 studies -->

These approaches are fundamentally limited by the available data. Large-scale replications are arduous and expensive to run, so only a few large-scale replication datasets exist, and most analyses analyses draw heavily on same small set of data points. In particular, the RP:P dataset is much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016;@anderson2016] -- to the point that much of what we think we know about replicability may be over-fit to the 100 studies included in RP:P. More replication data is needed to identify markers of replicability and targets for potential future experimental approaches.

Further, none of these studies focused on features of experimental design, such as within-subjects designs or the use of repeated measures. There has been speculation that both of these factors should be linked to increased replicability due to their role in enabling increased statistical precision. Within-subjects designs lead to more precise experimental estimates by allowing the estimation of correlated person-level variation across conditions; repeated measures allow for more precise estimation by averaging out measurement error. Both are often recommended by methodologists as part of good measurement practices, at least when they are feasible  [@rosenthal2008;@greenwald1976; @frank2023].

@@

@@

These replications were all conducted on short time scales, within a constrained class budget. In some cases the budget limited the number of participants who could be recruited, occasionally below what the original study included or what power analyses suggested. In nearly all cases, replications were conducted online, with recruitment from Amazon Mechanical Turk (the default from 2011 to 2020) or Prolific (the default from 2021 -- 2022).

@@

```{r smd, out.height="25%", fig.width=6, fig.height=3, fig.pos="ht", fig.cap="Relationship between effect size of the original study, effect size of the replication study, and subjective replication success rating, for those studies where effect size was applicable."}
d |> 
  filter(target_d_calc < 5, rep_d_calc < 5) |>
  ggplot(aes(x=target_d_calc, y=rep_d_calc))+

@@

@@

Figure \ref{fig:smd} shows the relationship between original standardized effect size, replication effect size, and subjective replication score. Some studies replicated withed similar size effects to the original, and others failed to replicate, with replication effect sizes near zero. On average, there was a diminution of effect sizes from original to replication. This pattern of results is consistent with the results of RP:P [@openscienceconsortium2015].

@@

Some multi-site replication projects have found heterogeneity in effect sizes across replication sites [@ebersole2020; @klein2018; @olsson2020]. If we assume that the level of heterogeneity in hypothetical multi-site replications of these effects is the same as the average level heterogeneity found in prior multi-site replications in psychology [$\tau=.21$ in units of standardized mean difference; @olsson2020], then `r round(tau_pct,1)`% (`r tau_predIntcount`/`r tau_predInttotal`) of replication effect sizes are distributionally consistent with the original effect size. However, more work on understanding heterogeneity is needed to understand what levels of heterogeneity to expect across different implementations of the same experiment, and how considerations of heterogeneity should impact interpretations of both novel results and replication results.

@@

cors <- tibble(preds = preds) |>
  mutate(r = sapply(preds, function(x) sub_cor(x, stat = "estimate")),
         p = sapply(preds, function(x) sub_cor(x, stat = "p"))) |> 
  mutate(Predictors=factor(preds, levels=c("social", "other_psych", "non_psych", "is_within", "single_vignette", "change_platform", "open_data", "open_mat", "stanford", "pub_year", "log_trials", "log_sample", "log_ratio_ss"), labels=c("Social", "Other psych", "Non psych", "Within subjects", "Single vignette", "Switch to online", "Open data", "Open materials", "Stanford", "Publication year", "Log n trials", "Log original sample size", "Log rep/orig sample"))) |> arrange(desc(r)) |> select(Predictors, r, p)

@@

Many predictors individually correlated with subjective replication success (Table \ref{tab:cor}). Predictors of higher replicability included within-subjects designs, larger numbers of trials per subject, and the original study having openly accessible data. Predictors of lower replicability included single-vignette studies (those with only one experimental stimulus per condition), classification as social psychology, and study pairs where the original study was in-person and the replication switched to online.

Distributions of study outcomes across some of these properties are shown in Figure \ref{fig:predictors-graph}. To quantify the effects of these single predictors, we ran logistic models predicting the subjective replication score (on a 0--1 scale) as a function of the single predictor, and exponentiated the coefficient to obtain an odds ratio. Both social and cognitive psychology studies were well represented in the replication sample, and the cognitive psychology studies replicated `r round(1/subfield, 2)` times as often as social psychology studies. Within and between subjects designs were both common, and within-subjects designs replicated `r within |> round(2)` times as much. Studies with multiple vignettes replicated `r round(1/vignette, 2)` times as often as single vignetted studies. However, there were strong correlations among these experimental features as well as between these experimental features and specific subfields.

Studies with open data, which almost always also had open materials as well, tended to replicate more than studies without open data. Nearly all replications studies were conducted online, but original studies were split between using in-person and online recruitment. Replications that switched to online were less likely to replicate than those that had the same modality as the original (generally both online, in a few cases both in-person).  While online studies in general show comparable results to studies conducted in person [@crump2013], switching the modality does decrease the closeness of the replication, and some studies done in person may not have been well-adapted (e.g., inductions might have been weaker or instructions might have been insufficient). These factors -- open materials, open data, and online samples -- are more common in more recent studies, and so these effects may partially reflect temporal trends.

@@

within <- dat |> mutate(is_within=ifelse(is_within, "Within \nsubjects", "Between\nsubjects")) |> 
  count(is_within, sub_rep, .drop=F) |> ggplot(aes(x=is_within, y=n, fill=sub_rep))+
  geom_bar(stat="identity", position="dodge")+
  scale_fill_viridis(discrete=T)+
  coord_cartesian(ylim=c(0,55), expand=F)+
  theme(axis.title.y=element_blank(), axis.title.x=element_blank(), legend.position="none")

@@

While a number of features show individual correlations with the subjective replication score, many correlate with one another. To determine which predictors were the strongest, we ran a series of pre-registered Bayesian regression models regularized using a horseshoe shrinkage prior [@carvalho09]. We first ran models using all of the original-replication pairs (N=176), but without original effect size and original $p$ value as predictors, as they were uncodable for some pairs. We next ran models including all predictors, but on only the subset of data where all predictors were available (N=136).

Coefficient estimates from the two models predicting the subjective replication scores are shown in Figure \ref{fig:mod-results}; see Supplement for full model results. Due to a large number of predictors coupled with a small and noisy dataset, there is much uncertainty around the coefficients even with strong regularization. The general directions of coefficients are consistent with the effects of the predictors in isolation.

Within-subjects designs stand out as the strongest correlate of replicability in the model with all the data (`r within_1`). In the model with all predictors, but less data, within-subjects designs remain strong (`r within_3`), and standardized effect size is also substantially related to subjective replication score (`r smd_3`). Both effects are robust to a sensitivity analysis including only studies with close replications and matching statistical tests (within-subjects `r within_sens`; effect size `r smd_sens`).

We also ran models predicting our secondary outcome measures: whether the replication effect was within the prediction interval of the original effect and what the p-original was between the replication and original. Both these models had more uncertain estimates. While the credible intervals were wide, the general patterns of predictor direction and relative strength were similar to the subjective replication models (see Supplement). The strongest predictors for prediction interval were still within-subjects designs (`r within_2`) and studies with larger effect sizes (`r smd_2`).

@@

do_draws <- function(model){
  draws <- model |> gather_draws(`b_[a-zA-Z_(): ]+`, regex=TRUE) |> 
    mutate(Term=str_sub(.variable, 3,-1) |> str_replace("M","_"), Estimate=.value) |>  filter(Term!="Intercept") |> 
    mutate(Term=factor(Term, levels=c("subfieldsocial", "subfieldother_psych", "subfieldnon_psych", "is_within", "single_vignette", "change_platform", "open_data", "open_mat", "stanford", "z_pub_year", "z_log_trials", "z_log_sample", "z_log_ratio_ss","z_log_p","z_target_d_calc"), labels=c("Social", "Other psych", "Non psych", "Within subjects", "Single vignette", "Switch to online", "Open data", "Open materials", "Stanford", "Publication year", "Log n trials", "Log original sample size", "Log rep/orig sample", "P-value", "Original effect size"))) 
}

@@

Here, we take a functional approach to assessing replicability, framing both our methods and interpretation around the idea of whether work can be repeated by an early-career scientist. We took advantage of 11 years of graduate student replication projects to study both the overall level of replication success and the correlational predictors of replication in a previously-unused dataset. In line with previous results, we found a `r round(pluck(sub_rep_rate,1))`% replication rate, with some studies showing effect sizes similar to the original and others much smaller. Within-subjects designs, work in the subfield of cognitive psychology, and the original and replication both using online samples stood out as the strongest correlates of replication success. As many of these predictors interrelate with one another, we ran regularized regressions with all the predictors at once. Due to our small sample, model estimates were uncertain, but within-subjects designs and large original effect sizes were the strongest predictors.

@@

Our results are limited by the sample of studies we included, which are limited in number and may not be representative of the studies of interest to psychologists as a whole. Further, our predictor variables were not manipulated, so they cannot be interpreted as causing (non-)replication, but only as correlational markers. Some of the correlates are most easily interpreted as being about the original study, and others reflect the closeness of the replication to the original. For instance, while within-subjects designs are more likely to replicate than between-subjects designs, this predictor could also be related to the types of experiments that tend to be run in each design. Given the strong relationship of this factor to replicability, slightly more skepticism and critical reading of between-subjects designs may be warranted, but this correlation, by itself, does not mean scientists should prefer within-subjects designs.

@@

The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 254) a graduate-level experimental methods class taught at Stanford by MCF from 2011 to 2022. This class is commonly taken by first year graduate students in psychology and related disciplines, and it has been a requirement of the Psychology PhD since around 2015. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include an example as well as the template given to students at TODO example and template.

@@

Four of these replication projects were included in RP:P, and 10 of them were previously reported in @hawkins2018.

@@

This subjective replication outcome was chosen because it already existed, could be applied to all projects (regardless of type and detail of statistical reporting), and did not rely solely on one statistical measure. As a complement, we also identified a "key" statistical test for each paper (see below for details), and if possible, computed p-original and prediction interval at this statistic, following @mathur2020. p-original was a continuous measure of the p-value on the hypothesis that the original and replication samples come from the same distribution. Prediction interval was a binary measure of whether the replication outcome fell within the prediction interval of the original outcome measure.

@@

We coded the number of subjects, post-exclusions. We coded whether a study had a between-subjects, within-subjects, or mixed design; for the analysis, mixed studies were counted as within-subjects designs. In the analysis, we used a log-scale for number of subjects and numbers of trials.

@@

### Determination and coding of key statistical measure
For each study pair, we used one key measure of interest for which we calculated the predictor variables of p-value and effect size and the statistical outcome measures p_original and prediction interval. If the student specified a single key measure of interest and this was a measure that was reported in both the original paper and replication, we used that measure. If a student specified multiple, equally important, key measures, we used the first one. When students were not explicit about a key measure, we used other parts of their report (including introduction and power analysis) to determine what effect and therefore what result they considered key. In a few cases, we went back to the original paper to find what effect was considered crucial by the original authors. When the measures reported by the student did not cleanly match their explicit or implicitly stated key measure, we picked the most important (or first) of the measures that were reported in both the original and replication. These decisions could be somewhat subjective but importantly they were made without reference to replication outcomes.

@@

We then calculated p values, effect sizes, p-original, and prediction intervals. We choose to recalculate p values and effect sizes from the means or test statistic rather than use reported measures when possible because we thought this would be more reliable and transparent. The means and test statistics are more likely to have been outputted programmatically and copied directly into the text. In contrast, p-values are often reported as <.001 rather than as a point value, and effect size derivations may be error prone. By recording the raw statistics we used and using our available code to calculate other measures, we are transparent, as the test statistics can be searched for in the papers, and all processing is documented in code.

In some cases, p-values or effect sizes were not calculable either due to insufficient reporting (ex. reporting a p-value but no other statistics from a test) or key measures where p-values and effect sizes did not apply (ex. PCA as measure of interest). Where studies reported beta estimates and standard errors or proportions, standardized effects sizes are not an applicable measure, but we were still able to calculate p-original and prediction interval.

@@

In regression analyses, we used standardized mean difference and log p-value as predictors.

@@

TODO citation around horseshoe prior

@@

We ran a model predicting the subjective replication score on the basis of demographic and experimental predictors on the entire dataset. We ran two models predicting p_original and prediction interval from demographic and experimental predictors on the subset of data where we had p_original and prediction intervals. Then, on the smaller subset of the data where we had effect sizes and p-values, we re-ran these three models with those as additional predictor variables.

The subjective replication scores were coded on [0, .25, .5, .75, 1], and we ramapped these to 1-5 to run an ordinal regression predicting replication score. We ran logistic regressions predicting prediction interval and linear regressions predicting p_original.

@@

Acknowledge people here. `{-}` useful to not number this section.

>>>>

Cumulative scientific progress requires empirical results that are robust enough to support theory construction and extension. Yet in psychology, some prominent findings have failed to replicate, and large-scale studies suggest replicability issues are widespread. The identification of predictors of replication success is limited by the difficulty of conducting large samples of independent replication experiments, however: most investigations re-analyse the same set of ~170 replications. We introduce a new dataset of 176 replications from students in a graduate-level methods course. Replication results were judged to be successful in 49% of replications; of the 136 where comparable effect sizes could be computed, 46% had point estimates within the prediction interval of the original outcome. Larger original effect sizes and within-participants designs were especially related to replication success. Our results indicate that, consistent with prior reports, the robustness of the psychology literature is low enough to limit cumulative progress by student investigators.
:::

@@

A few large-scale investigations have measured replication rates in samples of psychology studies. The first of these, RP:P, sampled roughly 100 studies from articles published in three top psychology journals in 2008 and distributed the studies across participating labs, finding an overall replication rate of around 40% [@openscienceconsortium2015]. The followup Many Labs studies investigated heterogeneity using short target studies that each compared two conditions. These study designs were not representative of the psychology literature as a whole, and due to the goal of measuring heterogeneity, they had large overall samples across multiple sites. Across Many Labs 1--3, only 29 of 51 target effects (57%) replicated [@klein2014; @klein2018; @ebersole2016]. @camerer2018 included all 21 behavioral social science studies from Nature and Science from 2010-2015 that were feasible to replicate. They consulted original authors and had high power to detect even small effects; under these conditions and with this sample, the replication rate was around 60%.

@@

Further, the methods used in these studies are not sufficient to yield an unbiased estimate of this quantity. In no case were studies randomly sampled from the literature; instead replication projects sampled from specific journals and adjusted the sample for feasibility concerns. These reasonable decisions further undermine the interpretation of the results from these studies as representing the proportion of true findings in the psychology literature as a whole. 
<!-- Sampling at random from the literature may not even be desirable, as arguably a literature will succeed if useful discoveries come out of it, not if random findings are true [@wilson2020].  -->

Rather than aim for some measure of true replicability, perhaps we should contextualize the true estimand for replication efforts based on their methodologies and outcome measures. Through this lens, RP:P estimated the rate at which findings from relatively simple experiments published in a few well-known journals at a particular time could be replicated in a typical psychology lab. Many Labs estimated the rate at which well-known, two-condition findings replicate in very large samples. @camerer2018 estimated which prestigious journal findings replicate when conducted in a highly-resourced environment with expert involvement. All of these could be potentially desirable estimands.

But in practice, most scientific work is conducted by graduate students with limited time, limited budgets, and limited access to experts. If students cannot replicate a finding under these circumstances, they cannot build on it in their own empirical or theoretical work. How replicable are findings in the literature for graduate students operating under these less-than-ideal conditions? We address this question here.

@@

In search of such features, RP:P examined correlates of replicability in the RP:P sample and found that studies in cognitive psychology (as opposed to social psychology) and studies with larger effect sizes and smaller $p$ values were more likely to replicate. Using these same data combined with a few other smaller samples, @altmejd2019 examined statistical and demographic features of replication studies and identified larger sample sizes, larger effect sizes, and simple effects (as opposed to interaction terms) as predictive of replication.^[While most approaches to correlates of replicability have been correlational, experimental approaches can be used to test potential interventions [@protzko2020;@ebersole2020]. Experiments can be very valuable as tests of specific causes of non-replication, but they are expensive and time-consuming to conduct, and not all factors affecting replication success can be manipulated experimentally.]
<!-- showed, across 16 studies, that better methodological practices led to replication rates that matched theoretical expectations based on sampling error, with replication effect sizes comparable with the original. Many Labs 5 added expert (original author) advice to a replication process, and found that it did not substantially increase the replication rate for the 10 studies -->

These approaches are fundamentally limited by the available data. Large-scale replications are arduous and expensive to run, so only a few large-scale replication datasets exist, and most analyses draw heavily on same small set of data points. In particular, the RP:P dataset is much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016;@anderson2016] -- to the point that much of what we think we know about replicability may be over-fit to the 100 studies included in RP:P.

Further, none of these studies focused on features of experimental design, such as within-participants designs or the use of repeated measures. There has been speculation that both of these factors should be linked to increased replicability due to their role in enabling increased statistical precision. Within-participants designs lead to more precise experimental estimates by allowing the estimation of correlated person-level variation across conditions; repeated measures allow for more precise estimation by averaging out measurement error. Both are often recommended by methodologists as part of good measurement practices, at least when they are feasible  [@rosenthal2008;@greenwald1976; @frank2023].

@@

target_dist <- quantile(d$target_N)
rep_dist <- quantile(d$replication_N)
ratio_dist <- d |> mutate(ratio_N=replication_N/target_N) |> pull(ratio_N) |> quantile()

@@

These replications were all conducted on short time scales, within a constrained class budget. In some cases the budget limited the number of participants who could be recruited, occasionally below what the original study included or what power analyses suggested. The replications had a median post-exclusion sample size that was `r round(ratio_dist[3]*100)`% of the original sample size (interquartile range `r round(ratio_dist[2]*100)`%-`r round(ratio_dist[4]*100)`%)^[Replications had a median of `r round(rep_dist[3])` participants (interquartile range: `r round(rep_dist[2])` - `r round(rep_dist[4])`), while original studies had a median of `r round(target_dist[3])` participants (interquartile range: `r round(target_dist[2])` - `r round(target_dist[4])`).].  In nearly all cases, replications were conducted online, with recruitment from Amazon Mechanical Turk (the default from 2011 to 2020) or Prolific (the default from 2021 -- 2022).

@@

```{r smd, out.height="25%", fig.width=6, fig.height=3, fig.pos="ht", fig.cap="Relationship between effect size of the original study, effect size of the replication study, and subjective replication success rating, for those studies where effect size was applicable (N=112)."}
d |> 
  filter(target_d_calc < 5, rep_d_calc < 5) |>
  ggplot(aes(x=target_d_calc, y=rep_d_calc))+

@@

# sample size considerations

# ggplot(d, aes(x=target_N, y=replication_N, color=as.factor(sub_rep)))+geom_point()+coord_trans(x="log", y="log")+geom_abline(slope=1)+ scale_color_viridis(discrete=T, direction = 1, 
#                       name = "Subjective\nReplication\nSuccess\n")
# d |> mutate(bigger_orig=replication_N<.9*target_N,
#             bigger_rep=replication_N>1.1*target_N,
#             similar=!bigger_orig & !bigger_rep) |> 
#   group_by(bigger_orig, bigger_rep, similar) |> tally()

@@

Figure \ref{fig:smd} shows the relationship between original standardized effect size, replication effect size, and subjective replication score. Some studies replicated with similar size effects to the original, and others failed to replicate, with replication effect sizes near zero. On average, there was a diminution of effect sizes from original to replication. This pattern of results is consistent with the results of RP:P [@openscienceconsortium2015].

@@

Some multi-site replication projects have found heterogeneity in effect sizes across replication sites [@ebersole2020; @klein2018; @olsson2020]. As a test of sensitivity to heterogeneity, we assumed that the level of heterogeneity in hypothetical multi-site replications of our sampled articles was the same as the average level heterogeneity found by @olssen2020 in prior multi-site replications in psychology ($\tau=.21$ in units of standardized mean difference). Under this assumption, `r round(tau_pct)`% (`r tau_predIntcount`/`r tau_predInttotal`) of replication effect sizes are distributionally consistent with the original effect size. However, more work on understanding heterogeneity is needed to understand what levels of heterogeneity to expect across different implementations of the same experiment, and how considerations of heterogeneity should impact interpretations of both novel results and replication results.

@@

cors <- tibble(preds = preds) |>
  mutate(r = sapply(preds, function(x) sub_cor(x, stat = "estimate")),
         p = sapply(preds, function(x) sub_cor(x, stat = "p"))) |> 
  mutate(Predictors=factor(preds, levels=c("social", "other_psych", "non_psych", "is_within", "single_vignette", "change_platform", "open_data", "open_mat", "stanford", "pub_year", "log_trials", "log_sample", "log_ratio_ss"), labels=c("Social", "Other psych", "Non psych", "Within participants", "Single vignette", "Switch to online", "Open data", "Open materials", "Stanford", "Publication year", "Log n trials", "Log original sample size", "Log rep/orig sample"))) |> arrange(desc(r)) |> select(Predictors, r, p)

@@

Many predictors individually correlated with subjective replication success (Table \ref{tab:cor}). Predictors of higher replicability included within-participants designs, larger numbers of trials per participant, and the original study having openly accessible data. Predictors of lower replicability included single-vignette studies (those with only one experimental stimulus per condition), classification as social psychology, and study pairs where the original study was in-person and the replication switched to online.

Distributions of study outcomes across some of these properties are shown in Figure \ref{fig:predictors-graph}. To quantify the effects of these single predictors, we ran logistic models predicting the subjective replication score (on a 0--1 scale) as a function of the single predictor, and exponentiated the coefficient to obtain an odds ratio. Both social and cognitive psychology studies were well represented in the replication sample, and the cognitive psychology studies replicated `r round(1/subfield, 2)` times as often as social psychology studies. Within and between participants designs were both common, and within-participants designs replicated `r within |> round(2)` times as much as between-participants designs. Studies with multiple vignettes replicated `r round(1/vignette, 2)` times as often as single vignetted studies. However, there were strong correlations among these experimental features as well as between these experimental features and specific subfields.

Studies with open data, which almost always also had open materials as well, tended to replicate more than studies without open data. Nearly all replication studies were conducted online, but original studies were split between using in-person and online recruitment. Replications that switched to online were less likely to replicate than those that had the same modality as the original (generally both online, in a few cases both in-person).  While online studies in general show comparable results to studies conducted in person [@crump2013], switching the modality does decrease the closeness of the replication, and some studies done in person may not have been well-adapted (e.g., inductions might have been weaker or instructions might have been insufficient). These factors -- open materials, open data, and online samples -- are more common in more recent studies, and so these effects may partially reflect temporal trends.

@@

within <- dat |> mutate(is_within=ifelse(is_within, "Within \nparticipants", "Between\nparticipants")) |> 
  count(is_within, sub_rep, .drop=F) |> ggplot(aes(x=is_within, y=n, fill=sub_rep))+
  geom_bar(stat="identity", position="dodge")+
  scale_fill_viridis(discrete=T)+
  coord_cartesian(ylim=c(0,55), expand=F)+
  theme(axis.title.y=element_blank(), axis.title.x=element_blank(), legend.position="none")

@@

While a number of features show individual correlations with the subjective replication score, many correlate with one another. To determine which predictors were the strongest, we ran a series of pre-registered Bayesian regression models regularized using a horseshoe shrinkage prior [@carvalho09]. We first ran models using all of the original-replication pairs (N=176), but without original effect size and original $p$ value as predictors, as they were uncodable for some pairs. We next ran models including all predictors, but on only the subset of data where all predictors were available (N=112).

Coefficient estimates from the two models predicting the subjective replication scores are shown in Figure \ref{fig:mod-results}. Due to a large number of predictors coupled with a small and noisy dataset, there is much uncertainty around the coefficients even with strong regularization. The general directions of coefficients are consistent with the effects of the predictors in isolation.

Within-participants designs stand out as the strongest correlate of replicability in the model with all the data (`r within_1`). In the model with all predictors, but less data, within-participants designs remain strong (`r within_3`), and standardized effect size is also substantially related to subjective replication score (`r smd_3`). Both effects are robust to a sensitivity analysis including only studies with close replications and matching statistical tests (within-participants `r within_sens`; effect size `r smd_sens`).

We also ran models predicting our secondary outcome measures: whether the replication effect was within the prediction interval of the original effect and what the $p$-original was between the replication and original. Both these models had more uncertain estimates. While the credible intervals were wide, the general patterns of predictor direction and relative strength were similar to the subjective replication models (see Supplement for coefficients from all models). The strongest predictors for prediction interval were still within-participants designs (`r within_2`) and studies with larger effect sizes (`r smd_2`).

@@

do_draws <- function(model){
  draws <- model |> gather_draws(`b_[a-zA-Z_(): ]+`, regex=TRUE) |> 
    mutate(Term=str_sub(.variable, 3,-1) |> str_replace("M","_"), Estimate=.value) |>  filter(Term!="Intercept") |> 
    mutate(Term=factor(Term, levels=c("subfieldsocial", "subfieldother_psych", "subfieldnon_psych", "is_within", "single_vignette", "change_platform", "open_data", "open_mat", "stanford", "z_pub_year", "z_log_trials", "z_log_sample", "z_log_ratio_ss","z_log_p","z_target_d_calc"), labels=c("Social", "Other psych", "Non psych", "Within participants", "Single vignette", "Switch to online", "Open data", "Open materials", "Stanford", "Publication year", "Log n trials", "Log original sample size", "Log rep/orig sample", "P-value", "Original effect size"))) 
}

@@

Here, we take a functional approach to assessing replicability, framing both our methods and interpretation around the idea of whether work can be repeated by an early-career scientist. We took advantage of 11 years of graduate student replication projects to study both the overall level of replication success and the correlational predictors of replication in a previously-unused dataset. In line with previous results, we found a `r round(pluck(sub_rep_rate,1))`% replication rate, with some studies showing effect sizes similar to the original and others much smaller. Within-participants designs, work in the subfield of cognitive psychology, and the original and replication both using online samples stood out as the strongest correlates of replication success. As many of these predictors interrelate with one another, we ran regularized regressions with all the predictors at once. Due to our small sample, model estimates were uncertain, but within-participants designs and large original effect sizes were the strongest predictors.

@@

Our results are limited by the sample of studies we included, which are limited in number and may not be representative of the studies of interest to psychologists as a whole. Further, our predictor variables were not manipulated, so they cannot be interpreted as causing (non-)replication, but only as correlational markers. Some of the correlates are most easily interpreted as being about the original study, and others reflect the closeness of the replication to the original. For instance, while within-participants designs are more likely to replicate than between-participants designs, this predictor could also be related to the types of experiments that tend to be run in each design. Given the strong relationship of this factor to replicability, slightly more skepticism and critical reading of between-participants designs may be warranted, but this correlation, by itself, does not mean scientists should prefer within-participants designs.

@@

The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 254) a graduate-level experimental methods class taught at Stanford by MCF from 2011 to 2022. This class is commonly taken by first year graduate students in psychology and related disciplines, and it has been a requirement of the Psychology PhD since around 2015. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include some example reports we have permission to share and the template given to students at TOOD OSF REPO.

@@

Four of the replication projects were included in RP:P, and 10 were previously reported in @hawkins2018 (which reported 11 student replications from the 2015-2016 class, one of those was excluded from the current sample for being non-experimental).

@@

This subjective replication outcome was chosen because it already existed, could be applied to all projects (regardless of type and detail of statistical reporting), and did not rely solely on one statistical measure. As a complement, we also identified a "key" statistical test for each paper (see below for details), and if possible, computed $p$-original and prediction interval at this statistic, following @mathur2020. $p$-original was a continuous measure of the $p$ value on the hypothesis that the original and replication samples come from the same distribution. Prediction interval was a binary measure of whether the replication outcome fell within the prediction interval of the original outcome measure.

@@

We coded the number of participants, post-exclusions. We coded whether a study had a between-participants, within-participants, or mixed design; for the analysis, mixed studies were counted as within-participants designs. In the analysis, we used a log-scale for number of participants and numbers of trials.

@@

### Determination and coding of key statistical measure
For each study pair, we used one key measure of interest for which we calculated the predictor variables of $p$ value and effect size and the statistical outcome measures $p$-original and prediction interval. If the student specified a single key measure of interest and this was a measure that was reported in both the original paper and replication, we used that measure. If a student specified multiple, equally important, key measures, we used the first one. When students were not explicit about a key measure, we used other parts of their report (including introduction and power analysis) to determine what effect and therefore what result they considered key. In a few cases, we went back to the original paper to find what effect was considered crucial by the original authors. When the measures reported by the student did not cleanly match their explicit or implicitly stated key measure, we picked the most important (or first) of the measures that were reported in both the original and replication. These decisions could be somewhat subjective but importantly they were made without reference to replication outcomes.

@@

We then calculated $p$ values, effect sizes, $p$-original, and prediction intervals. We choose to recalculate p values and effect sizes from the means or test statistic rather than use reported measures when possible because we thought this would be more reliable and transparent. The means and test statistics are more likely to have been outputted programmatically and copied directly into the text. In contrast, $p$ values are often reported as <.001 rather than as a point value, and effect size derivations may be error prone. By recording the raw statistics we used and using our available code to calculate other measures, we are transparent, as the test statistics can be searched for in the papers, and all processing is documented in code.

In some cases, $p$ values or effect sizes were not calculable either due to insufficient reporting (ex. reporting a $p$ value but no other statistics from a test) or key measures where $p$ values and effect sizes did not apply (ex. PCA as measure of interest). Where studies reported beta estimates and standard errors or proportions, standardized effects sizes are not an applicable measure, but we were still able to calculate $p$-original and prediction interval.

@@

In regression analyses, we used standardized mean difference and log $p$ value as predictors.

@@

@@

We ran a model predicting the subjective replication score on the basis of demographic and experimental predictors on the entire dataset. We ran two models predicting $p$-original and prediction interval from demographic and experimental predictors on the subset of data where we had $p$-original and prediction intervals. Then, on the smaller subset of the data where we had effect sizes and $p$ values, we re-ran these three models with those as additional predictor variables.

The subjective replication scores were coded on [0, .25, .5, .75, 1], and we ramapped these to 1-5 to run an ordinal regression predicting replication score. We ran logistic regressions predicting prediction interval and linear regressions predicting $p$-original.

@@

TODO We are grateful to the students in PSYCH 254 and PSYCH 251 for conducting the replication projects included here. We thank Ben Prystawski and the FriSem audience for feedback on earlier versions of this work. TODO GRANTS??
